Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-03-31 manuscript/supplementaryRW.Rmd
update supplementary

# Supplementary Material 1: Protocol

## General steps

Each modeling team will run their simulations prior to the workshop, and
will report on the results before ([short written report](#report)) and
during the workshop. The general approach for the workshop is:

1.  Each team sets up their model to simulate the regeneration niche of
    [eleven major European tree species](#speciescomposition) within the
    known, defined range of environmental conditions of 200 sites. They
    will run at least one simulation per site. It is important to
    conduct mixed-species simulations, not series of monospecific
    stands.

2.  The model is set up for each site based on the [input variable
    information](#inputvariables) provided in this protocol, and the
    simulations need to be run until the [equilibrium is
    reached](#equilibrium).

3.  Each team reports on the results of these exercises in terms of a
    [standardized output dataset](#outputdata) (including density
    variables per 10-yr interval and per ha, for each simulated site and
    species) and in the form of a brief [working report](#report) (\~4
    pages).

4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
    the regeneration niche as it results from each model and compare it
    to the empirical estimates of the environmental limits of the
    regeneration niche that will be made available to the participants
    shortly before the workshop. As a group effort, the projections of
    the niche properties arising from the different models will be
    compared. The exact details on how this will be done will be
    provided once we know how many modeling teams will contribute
    results.

## Input variables {#inputvariables}

In this section, we are providing a description of all the necessary
driver information to run the simulations (e.g., location, climate,
soil).

### Coordinates

We intend to provide the site information via unique site IDs. If actual
coordinates are needed, we can provide blurred site coordinates upon
reasonable request.

We provide time series of climatic variables from 1981 to 2018 in
hourly, daily and monthly resolution. Based on our survey the climate
data should be sufficient to run all models. Some variables such as
relative humidity or vapor pressure deficit must be calculated from the
variables we provide.

Note that the hourly time series has the lowest spatial resolution, thus
the daily time series should be used for spatial accuracy. If you need
hourly time series we recommend to use the daily time series and only
use the hourly time series to correct for hourly variation. For this
purpose we also provide hourly averages over the period 1981 to 2018 for
each hour of each day of the year. Only for calculating these averages
the 29th of February in leap years was removed from the time series to
obtain values for 365 days.

#### Data sources {#climsources}

The hourly time series are based on the Era5-land data set. Daily and
monthly time series are mostly based on CHELSA data which has a higher
spatial resolution. More details are given below or in the documentation
of the original data source.

The variable names are kept exactly like in the documentation of the
Era5 and CHELSA data sets to facilitate the use of the original data
documentation. Please don't be confused by the ending `*_hourly` in
`surface_thermal_radiation_downwards_hourly` and the two different
variable names for shortwave radiation: `rsds` for the values from
CHELSA and `surface_solar_radiation_downwards_hourly` for the values
from Era5.

All changes that were made to the original data are documented here:

-   The wind variable `wind_speed` is based on the u and v component of
    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
    variable `u_component_of_wind_10m` and $v$ is
    `v_component_of_wind_10m` of the Era5 data set.

-   The units of temperature variables `temperature_2m` (Era5), `tas`,
    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
    Celsius

-   The units of precipitation variables `total_precipitation_hourly`
    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
    respectively.

-   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
    to J.

We aggregated variables from hourly to daily to monthly values by using
means for temperature, surface pressure and wind speed and sums for
radiation and precipitation.

##### Era5-Land {#era5}

-   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
    present. Copernicus Climate Change Service (C3S) Climate Data Store
    (CDS). (23.01.2022),
    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
-   [Documentation of the hourly ERA5 data from the Google Earth
    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
-   Further information is available from the [official copernicus
    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).

##### CHELSA {#chelsa}

-   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
    with CHELSA v2.0. ISIMIP Repository.
    <https://doi.org/10.48364/ISIMIP.836809>

-   [Official CHELSA website](https://chelsa-climate.org)

#### Monthly time series 1981-2018

file name: monthly_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Daily time series 1981-2018

file name: daily_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Hourly time series 1981-2018

file name: hour_era5_200.zip

file name: yhourMean_era5_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average temperature (temperature_2m) [°C]

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   precipitation sum (total_precipitation_hourly) [mm]

-   surface solar (shortwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

-   surface thermal (longwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

file name: soil_quality_200.csv

**Data description**

The following variables are available:

-   Soil quality is given as a continuous value between 1 and 5.

**Data source**

-   The soil quality value is based on a random forest that was trained
    to predict expert knowledge based soil quality on a scale from 1 to
    5 from the [SoilGrids250 data
    set](https://www.isric.org/explore/soilgrids) and the [WISE data
    set](https://www.isric.org/explore/wise-databases). The most
    influential variables in the random forest were total depth to
    bedrock and sand content. Thus the variable is a good predictor for
    plant available water storage capacity.

## Topography and elevation

There was an error in a earlier version of this file. Please use the
link above to download the corrected file.

This is the old file and should not be used because the aspect column
contains elevation instead of the aspect.

file name: topo_data_200.csv

**Data description**

The following variables are available:

-   elevation [m.a.s.l]

-   slope [Degree]

-   aspect [Degree]

**Data source**

-   Topographic information was derived from the [ASTER DEM
    V3](https://lpdaac.usgs.gov/products/astgtmv003/).

-   Slope and aspect were calculated with QGIS. The specific tiles that
    were used are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
    and the processing steps in QGIS are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).

### Natural disturbances

The simulations must be run in the absence of natural disturbances.

## Management

The simulations must be run in the absence of management.

### Area simulated

From your model outputs, we would like to obtain samples on species
specific regeneration rates per decade and per ha. The modeling teams
decide on the simulated area and how they derive these samples. Please
describe the methodology used in the [output report](#report).

### Simulation duration and equilibrium {#equilibrium}

The simulation should be run from bare ground in the absence of
management to a simulated [equilibrium](#defequilibrium) ("potential
natural vegetation") with current climate. The goal is to derive the
[regeneration niche](#regenniche). The exact length of the simulation
should be decided by the modeling teams. Please describe the simulation
length used in your simulations in the [output report](#report).

[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
below). Both the width of the regeneration niche (i.e., in environmental
space) as well as the intensity of the recruitment process (i.e., the
number of ingrowth trees per area and per unit of time) are of interest.

[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
the "Potential Natural Vegetation".

### Species composition {#speciescomposition}

**Mixed species simulations** must be run (not multiple single-species
simulations).

**Option A: Species level**

The simulation should be run for **mixtures** of the following
species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*

Additional species can be included in the simulation if modelers think
this is important, but they should report only on these 11
species/genera because the benchmarking data refer to them only. Please
report the methodology used in the [output report](#report).

**Option B: PFTs**

If species-level simulation is not possible in your model, substitute
the species by PFTs and use the model's default approach. Please report
the methodology used in the [output report](#report). Explain as well
which PFTs represent which species for your model.

## Diameter threshold

The [output data](#outputdata) have to be provided for two diameter
thresholds: 7 and 10 cm.

## Output

We are expecting to get the number of recruits over a 10-yr interval per
species and per ha for each of the sites. However, we are also
interested in the **variation per site** (i.e., we would like to see the
range of estimates, not only the mean value per site). Each modeling
team will decide how to sample multiple values per site that will show
the variation in recruitment.

Specifically, you should provide us with multiple samples (n = 200; this
number is the same as the number of sites by sheer coincidence) of the
regeneration rate for spatial units of 1 ha (please do [not]{.underline}
provide average of multiple 1-ha samples). This can be done in various
ways, e.g., by sampling simulated data from the same 1-ha plot in the
equilibrium over time, or by sampling several 1-ha plots from the
simulated forest at one specific point in time (in the equilibrium); for
clarification, cf. three sample strategies proposed in [figures
1](#sample1), [2](#sample2) and [3](#sample3).

If, for some reason, you cannot submit **200 samples** of **1 ha**,
please **let us know** in the [workshop chat](#chat) so we can discuss
alternatives. The sampling from the simulations must happen **within**
the simulated phase of Potential Natural Vegetation (for more
information, see how [equilibrium](#defequilibrium), Potential [Natural
Vegetation, and regeneration niche](#regenniche) are defined in this
protocol).

![Figure 1. In this example there are 200 samples collected at year i at
equilibrium differentiated spatially with each square representing 1
ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}

![Figure 2. In this example, there are 100 spatially differentiated
samples taken in year i at equilibrium and 100 in year i+10. Each square
represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}

![Figure 3. In this example there is 1 spatially differentiated sample
taken in year i at equilibrium and remeasured every 10 years for 2000
years until i+2000. The square represents 1
ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}

### Output data {#outputdata}

Please provide the simulation output data using the following structure:

-   Each row of your output dataset contains the recruitment of 10-year
    interval per species and ha for a unique simulation site.

-   Each column of your output dataset contains one of these variables:

    -   `sample`: this will identify the different samples per site
        (1:n, with n $\ge$ 200)

    -   `site`: site identifier (1:200)

    -   `species`: species simulated with the following categories:

        -   fasy (for *Fagus sylvatica*)
        -   pabi (for *Picea abies*)
        -   abal (for *Abies alba*)
        -   cabe (for *Carpinus betulus*)
        -   tico (for *Tilia cordata*)
        -   acps (for *Acer pseudoplatanus*)
        -   betu (for *Betula* spp.)
        -   frex (for *Fraxinus excelsior*)
        -   quer (for *Quercus* spp.)
        -   algl (for *Alnus glutinosa*)
        -   pisy (for *Pinus sylvestris*)
        -   If you have some other categories please create your own
            species codes and provide the legend in the output report.
            We prefer a lumped category "othr".

    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
        decade.

    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
        and decade.

    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
        decade

    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
        or 10 cm).

### Output report {#report}

To guide the writing of the output report, please refer to the following
questions:

-   What climate data have you used, and how have you created your model
    climate input parameters?

-   How have you defined the species composition in your model
    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
    species categories used.

-   How have you defined the simulation area for your model (e.g., size
    and other relevant site aspects)?

-   What are the relevant aspects of the dispersal (if any) and seed
    input parameters in your model?

-   What is the simulation length, and how have you determined the
    simulation time until the equilibrium?

-   If you have run more than one simulation per site, please describe
    the differences of the approaches used.

-   Please describe how have you calculated each of the variables in the
    [output data](#outputdata).

### Uploading data and report

Please upload your data and output report as a **zip file** with the
name of your model and any other word that might help us to understand
the files, for example: *thenameofyourmodel_datareport.zip* or
*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.

You can upload the files here: LINK

# Supplementary Material 2: Model reports

@@

Yannek Käber Harald Bugmann

@@

SIBYLA is the simulator of forest biodynamics. It belongs to the
category of semi-empirical tree growth simulators. It consists of the
set of mathematical models and algorithms that are transformed into an
integrated software package SIBYLA Suite. The model has been developed
at the Department of Forest Management and Geodesy, Technical University
in Zvolen, Slovakia. The basis of the model was the modelling principle
and algorithms implemented in SILVA 2.2 (Pretzsch 1992, Kahn 1994). The
model is sensitive to climatic factors (length of growing season, mean
temperature during growing season, annual temperature amplitude, and
total precipitation during growing season). The climatic factors modify
height and diameter growth potential, and consequently tree increment.
The climatic factors also influence tree regeneration model. More
detailed information about Sibyla can be found in Fabrika (2005),
<http://etools.tuzvo.sk/sibyla/english/model.htm>.

@@

• What climate data have you used, and how have you created your model
climate input parameters?

@@

• Please describe how have you calculated each of the variables in the
output data

@@

Figure 1: Exemplary overview over output file.

@@

```{=tex}
\newpage
\beginsupplement
```
# Supplementary Material 3: Supplementary figures and tables

@@

>>>>

# Supplementary Material 1: Model reports

@@

Yannek Käber, Harald Bugmann

@@

SIBYLA is the simulator of forest biodynamics. It belongs to the
category of semi-empirical tree growth simulators. It consists of the
set of mathematical models and algorithms that are transformed into an
integrated software package SIBYLA Suite. The model has been developed
at the Department of Forest Management and Geodesy, Technical University
in Zvolen, Slovakia. The basis of the model was the modelling principle
and algorithms implemented in SILVA 2.2 (Pretzsch 1992, Kahn 1994). The
model is sensitive to climatic factors (length of growing season, mean
temperature during growing season, annual temperature amplitude, and
total precipitation during growing season). The climatic factors modify
height and diameter growth potential, and consequently tree increment.
The climatic factors also influence tree regeneration model.

More detailed information about Sibyla can be found in Fabrika (2005) and in the link: 
<http://etools.tuzvo.sk/sibyla/english/model.htm>.

@@

@@

@@

@@

```{=tex}
\newpage
\beginsupplement
```
# Supplementary Material 2: Supplementary figures and tables

@@

## References
